\section{Utility Functions Module (utils.py)}
\seclabel{utils}

The \texttt{utils.py} module provides a collection of utility functions that support the image enhancement process. These functions handle tasks such as parameter saving/loading, image comparison, metrics visualization, and format handling.

\subsection{Parameter Management}

\subsubsection{save\_params\_to\_json}
\begin{lstlisting}[language=Python]
def save_params_to_json(params: EnhancementParams, output_path: str) -> None
\end{lstlisting}

Saves enhancement parameters to a JSON file for later reuse.

\begin{itemize}
  \item \textbf{params}: An \texttt{EnhancementParams} object containing the enhancement settings
  \item \textbf{output\_path}: Path where the JSON file will be saved
\end{itemize}

\subsubsection{load\_params\_from\_json}
\begin{lstlisting}[language=Python]
def load_params_from_json(input_path: str) -> EnhancementParams
\end{lstlisting}

Loads enhancement parameters from a previously saved JSON file.

\begin{itemize}
  \item \textbf{input\_path}: Path to the JSON file containing parameters
  \item \textbf{Returns}: An \texttt{EnhancementParams} object with the loaded settings
\end{itemize}

\subsection{Image Comparison}

\subsubsection{create\_comparison\_image}
\begin{lstlisting}[language=Python]
def create_comparison_image(original: Image.Image, 
                           enhanced: Image.Image, 
                           params: Optional[EnhancementParams] = None,
                           title: str = "Image Enhancement Comparison") -> Image.Image
\end{lstlisting}

Creates a side-by-side comparison of the original and enhanced images, optionally displaying the enhancement parameters used.

\begin{itemize}
  \item \textbf{original}: The original PIL Image
  \item \textbf{enhanced}: The enhanced PIL Image
  \item \textbf{params}: Optional EnhancementParams to display on the comparison image
  \item \textbf{title}: Title for the comparison image
  \item \textbf{Returns}: A new PIL Image with the side-by-side comparison
\end{itemize}

This function is particularly useful for visually evaluating the effects of enhancement operations. The comparison image includes:
\begin{itemize}
  \item Original and enhanced images side by side
  \item Labels for each image
  \item Title at the top
  \item Enhancement parameters displayed at the bottom (if provided)
\end{itemize}

\subsection{Enhancement Analysis}

\subsubsection{plot\_enhancement\_metrics}
\begin{lstlisting}[language=Python]
def plot_enhancement_metrics(original_image: Union[str, Image.Image], 
                            enhanced_image: Union[str, Image.Image],
                            params: EnhancementParams,
                            show_plot: bool = True,
                            save_path: Optional[str] = None) -> None
\end{lstlisting}

Generates plots showing the effects of enhancement operations through histograms and visual comparison.

\begin{itemize}
  \item \textbf{original\_image}: Path to the original image or a PIL Image object
  \item \textbf{enhanced\_image}: Path to the enhanced image or a PIL Image object
  \item \textbf{params}: Enhancement parameters used
  \item \textbf{show\_plot}: Whether to display the plot interactively
  \item \textbf{save\_path}: Optional path to save the plot as an image file
\end{itemize}

The generated plot includes:
\begin{itemize}
  \item Original and enhanced images
  \item Grayscale histogram comparison
  \item RGB channel histograms comparison
  \item Summary of applied enhancement parameters
\end{itemize}

This function is valuable for quantitative analysis of enhancement effects and understanding how specific parameters affect image properties.

\subsection{Format Handling}

\subsubsection{get\_supported\_formats}
\begin{lstlisting}[language=Python]
def get_supported_formats() -> List[str]
\end{lstlisting}

Returns a list of supported image file formats for input and output operations.

\begin{itemize}
  \item \textbf{Returns}: List of supported image format extensions (e.g., '.jpg', '.png', etc.)
\end{itemize}

\subsection{Usage Examples}

\subsubsection{Saving and Loading Parameters}
\begin{lstlisting}[language=Python]
# Create enhancement parameters
params = EnhancementParams(brightness=1.2, contrast=1.3, sharpness=1.5)

# Save parameters to a file
save_params_to_json(params, "my_enhancement_params.json")

# Later, load the parameters
loaded_params = load_params_from_json("my_enhancement_params.json")
\end{lstlisting}

\subsubsection{Creating a Comparison Image}
\begin{lstlisting}[language=Python]
from PIL import Image
from src.img_enhance.enhancer import ImageEnhancer, EnhancementParams
from src.img_enhance.utils import create_comparison_image

# Load original image
original = Image.open("document.jpg")

# Create enhancer and parameters
params = EnhancementParams(brightness=1.2, contrast=1.5, binarize=True)
enhancer = ImageEnhancer(params)

# Enhance image
enhanced = enhancer.enhance(original)

# Create and save comparison
comparison = create_comparison_image(original, enhanced, params)
comparison.save("document_comparison.png")
\end{lstlisting}

\subsubsection{Analyzing Enhancement Effects}
\begin{lstlisting}[language=Python]
from src.img_enhance.utils import plot_enhancement_metrics
from src.img_enhance.enhancer import EnhancementParams

# Define parameters used
params = EnhancementParams(brightness=1.2, contrast=1.3, denoise=True)

# Generate and save analysis plot
plot_enhancement_metrics(
    "original.jpg",
    "enhanced.jpg",
    params,
    show_plot=False,
    save_path="enhancement_analysis.png"
)
\end{lstlisting} 